\textcolor{Navy}{\textsc{Convexity adjustments with a bit of Malliavin}}

### 2022-09-01 Latex/paper.tex
# New project test unitaries as examples

\begin{equation}
L_{E}(t_0,t_1,t_2) = \mathbb{E}^{\mathbb{Q}}_t\left(L_{E}(t_0,t_1,t_2)\right) + \int_{0}^{t_0} \mathbb{E}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) dW^{\mathbb{Q}}_s
\end{equation} 
therefore, if we take $\mathbb{E}^{\mathbb{Q}^{t2}}(\cdot)$ in the last equation, then we get
\begin{equation*}
CA(t, t_0, t_1, t_2) = - \mathbb{E}^{\mathbb{Q}^{t2}}\left(\int_{0}^{t_0} \mathbb{E}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) dW^{\mathbb{Q}}_s \right).
\end{equation*}
From (\ref{ois_forward_rate_curve}) and since $f_{ois}(t,T)$ is a $\mathbb{Q}^{T}$ martingale, we have that
\begin{equation*}
dW^{\mathbb{Q}^{t_2}} = dW^{\mathbb{Q}} + \nu(t,T) dt. 
\end{equation*}
Therefore, if we apply Girsanov's theorem to switch to measure $\mathbb{Q}^{t_2}$, we get that
\begin{equation}\label{ca_general_future}
CA(t, t_0, t_1, t_2) = \mathbb{E}^{\mathbb{Q}^{t2}}\left(\int_{0}^{t_0} \mathbb{E}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) \nu(s,t_2) ds \right)
\end{equation}
where $\nu(t,T)$ has been defined in (\ref{ois_forward_rate_curve}). Now, from the definition of $L_{E}(t,T)$ we have that
\begin{equation*}
D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)} D_s \left(\frac{P_{ois}(t_0,t_1)}{P_{ois}(t_0,t_2)}\right). 
\end{equation*}
If we use the representation formula (\ref{bond_ois}) we get that
\begin{equation*}
D_s \left(\frac{P_{ois}(t_0,t_1)}{P_{ois}(t_0,t_2)}\right) = \frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}
\end{equation*}
therefore
\begin{equation*}
D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}. 
\end{equation*}
From (\ref{short_rate_cheyette}) y we define $\beta(s,t, x, y) = \exp\left(-\int_{s}^{t} k_{u} \right) \eta(s, x, y)$ we have that
\begin{equation*}
D_s x_{t_0} = \beta(s,t_0, x_s, y_s)M(s,t_0)
\end{equation*}
where $M(s,t_0)=\exp\left(-\frac{\int_{s}^{t_0} \left(\partial_x \beta(u,t_0,x_s,y_s)\right)^{2} du}{2} + \int_{s}^{t_0} \partial_x \beta(u,t_0,x_s,y_s) dW^{\mathbb{Q}}_u \right)$. 
Then, we have that
\begin{align*}
D_s L_{E}(t_0,t_1,t_2) &= \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} \beta(s,t_0, x_s, y_s)M(s,t_0) \nonumber \\
=& \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) - G(t_0,t_1)\right)\beta(s,t_0, x_s, y_s)M(s,t_0).
\end{align*}
It is a common practice, freeze the state variables of the model to achieve some approximations. We can approximate $x_s$ and $y_s$ the next way
\begin{align*}
x_s &\approx x_0 \nonumber \\
y_s &\approx y_0 + \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0) du
\end{align*}
Therefore, 
\begin{equation}\label{approximation_clarkocone}
\mathbb{E}_s\left( D_s L_{E}(t_0,t_1,t_2) \right) = \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) - G(t_0,t_1)\right)\beta(s,t_0, x_0, \hat{y}_s)
\end{equation}
where $\hat{y}_s = y_0 + \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0) du$.
Now, from (\ref{ca_general_future}) and (\ref{approximation_clarkocone}) we have the next approximation for the convexity adjustmet for future

\hspace{1cm}
\fboxsep1em
\fbox{\parbox{350pt}
{
\begin{equation}\label{ca_approximation_futures}
CA(t,t_0,t_1) \approx \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) - G(t_0,t_1)\right) \int_{0}^{t_0}  \beta(s,t_0, x_0, \hat{y}_s) \nu(s,t_2) ds. 
\end{equation}
}}
\begin{example}
Let us to set 
\begin{align*}
g(T) &= \exp(-kT) \\
h(t) &= \sigma
\end{align*}
\end{example}
With this parametrization, the Cheyette model is equiavalent to Hull-White model. It is easy to show that under this parametrization, the convexity adjustment (\ref{ca_approximation_futures}) is
\begin{align*}
CA(t,t_0,t_1) & \approx \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(\exp(-k(t_1 - t_0)) - \exp(-k(t_1 - t_0)) \right) \sigma^{2} \exp(-k t_0)  \int_{0}^{t_0} \frac{\left(\exp(- k s) - \exp(- k t_2)   \right)}{k} ds.   
\end{align*}
In the next figure, we can check the accuracy of the last formula versus montecarlo. The parameters that we have used are $\sigma=0.05$, $k=0.003$ and flat curve with level $r=0.01$.

@@

\subsection{OIS futures}
In this we will derive the covexity adjustment for overnight-indexed swap (OIS) under. We will define for $t_0 < t_1$
\begin{align*}
R(t_0,t_1) &= \frac{\mathbb{E}_t^{Q}\left(\exp\left(\int_{t_0}^{t_1}r_{ois,u} du\right)\right) - 1}{t_1 - t_0} \\
R_{avg}(t_0,t_1) &= \frac{\int_{t_0}^{t_1}r_{ois,u} du}{t_1 - t_0}.
\end{align*}
We must observe, that both $R(\cdot,t_0,t_1)$ and $R_{avg}(\cdot,t_0,t_1)$  are not predictible and they are only observable in $t_1$. However, since
$R(\cdot,t_0,t_1)$ and $R_{avg}(\cdot,t_0,t_1)$ are flows that will be payed in $t_1$, we can considerer the expected value under the measure $\mathbb{Q}^{t_1}$ that will observable during whole period $[t_0, t_1]$. Let us define the next $\mathbb{Q}^{t_1}$ martingales  
\begin{align*} 
\bar{R}(t,t_0,t_1) &= \mathbb{E}^{\mathbb{Q}^{t_1}}\left( R(t_0,t_1)  \right) \\
\bar{R}_{avg}(t,t_0,t_1) &= \mathbb{E}^{\mathbb{Q}^{t_1}}\left( R_{avg}(t_0,t_1)  \right).
\end{align*}
We must do several obeservations. The first observation is that
\begin{align*}
F(t,t_0,t_1)=\mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R(t_0,t_1) \right) &= \frac{P_{ois}(t,t_0)- P_{ois}(t,t_1)}{t_1 - t_0}, \quad t \in [0,t_0] \\
F(t,t_0,t_1)=\mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R(t_0,t_1) \right) &= \frac{1}{t_1 - t_0} \left(\frac{\left( \exp\left(\int_{t}^{t_1}r_{ois,u} du\right)\right)}{P_{ois}(t_0,t)}-1\right), \quad  t \in  [t_0, t_1].
\end{align*}
The can define the convexity adjustment for $ R(t_0,t_1)$ the next way
\begin{equation}\label{R_ois_ca}
CA_{ois}(t,t_0,t_1) = F(t,t_0,t_1) - \bar{R}(t,t_0,t_1) 
\end{equation}
The second observation, we must do is that
\begin{equation*}
F_{avg}(t,t_0,t_1) = \mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R_{avg}(t_0,t_1) \right) =  
\end{equation*}

>>>>

\begin{equation}
L_{E}(t_0,t_1,t_2) = \mathbb{E}^{\mathbb{Q}}_t\left(L_{E}(t_0,t_1,t_2)\right) + \int_{0}^{t_0} \mathbb{E}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) dW^{\mathbb{Q}}_s
\end{equation} 
therefore, if we take $\mathbb{E}^{\mathbb{Q}^{t2}}(\cdot)$ in the last equation, then we get
\begin{equation*}
CA(t, t_0, t_1, t_2) = - \mathbb{E}^{\mathbb{Q}^{t2}}\left(\int_{0}^{t_0} \mathbb{E}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) dW^{\mathbb{Q}}_s \right).
\end{equation*}
From (\ref{ois_forward_rate_curve}) and since $f_{ois}(t,T)$ is a $\mathbb{Q}^{T}$ martingale, we have that
\begin{equation*}
dW^{\mathbb{Q}^{t_2}} = dW^{\mathbb{Q}} + \nu(t,T) dt. 
\end{equation*}
Therefore, if we apply Girsanov's theorem to switch to measure $\mathbb{Q}^{t_2}$, we get that
\begin{equation}\label{ca_general_future}
CA(t, t_0, t_1, t_2) = \mathbb{E}^{\mathbb{Q}^{t2}}\left(\int_{0}^{t_0} \mathbb{E}_{s}\left(D_s L_{E}(t_0,t_1,t_2) \right) \nu(s,t_2) ds \right)
\end{equation}
where $\nu(t,T)$ has been defined in (\ref{ois_forward_rate_curve}). Now, from the definition of $L_{E}(t,T)$ we have that
\begin{equation*}
D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)} D_s \left(\frac{P_{ois}(t_0,t_1)}{P_{ois}(t_0,t_2)}\right). 
\end{equation*}
If we use the representation formula (\ref{bond_ois}) we get that
\begin{equation*}
D_s \left(\frac{P_{ois}(t_0,t_1)}{P_{ois}(t_0,t_2)}\right) = \frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}
\end{equation*}
therefore
\begin{equation}\label{malliavin_derive_L}
D_s L_{E}(t_0,t_1,t_2) = \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} D_s x_{t_0}. 
\end{equation}
From (\ref{short_rate_cheyette}) y we define $\beta(s,t, x, y) = \exp\left(-\int_{s}^{t} k_{u} \right) \eta(s, x, y)$ we have that
\begin{equation}\label{malliavin_derive_x}
D_s x_{t_0} = \beta(s,t_0, x_s, y_s)M(s,t_0)
\end{equation}
where $M(s,t_0)=\exp\left(-\frac{\int_{s}^{t_0} \left(\partial_x \beta(u,t_0,x_s,y_s)\right)^{2} du}{2} + \int_{s}^{t_0} \partial_x \beta(u,t_0,x_s,y_s) dW^{\mathbb{Q}}_u \right)$. 
Then, we have that
\begin{align*}
D_s L_{E}(t_0,t_1,t_2) &= \frac{H(t_0,t_1)}{\delta_{t_1,t_2}H(t_0,t_2)}\frac{\left(\partial_{x}P_{ois}(t_0,t_1)P_{ois}(t_0,t_2) - \partial_{x}P_{ois}(t_0,t_2) P_{ois}(t_0,t_1) \right)}{P^{2}_{ois}(t_0,t_2)} \beta(s,t_0, x_s, y_s)M(s,t_0) \nonumber \\
=& \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) - G(t_0,t_1)\right)\beta(s,t_0, x_s, y_s)M(s,t_0).
\end{align*}
It is a common practice, freeze the state variables of the model to achieve some approximations. We can approximate $x_s$ and $y_s$ the next way
\begin{align*}
x_s &\approx x_0 \nonumber \\
y_s &\approx y_0 + \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0) du
\end{align*}
Therefore, 
\begin{equation}\label{approximation_clarkocone}
\mathbb{E}_s\left( D_s L_{E}(t_0,t_1,t_2) \right) = \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) - G(t_0,t_1)\right)\beta(s,t_0, x_0, \hat{y}_s)
\end{equation}
where $\hat{y}_s = y_0 + \int_{0}^{s} \exp\left(-2\int_{u}^{s} k_{u^{\prime}} du^{\prime} \right) \eta(u,x_0,y_0) du$.
Now, from (\ref{ca_general_future}) and (\ref{approximation_clarkocone}) we have the next approximation for the convexity adjustmet for future

\hspace{1cm}
\fboxsep1em
\fbox{\parbox{350pt}
{
\begin{equation}\label{ca_approximation_futures}
CA(t,t_0,t_1) \approx \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(G(t_0,t_2) - G(t_0,t_1)\right) \int_{0}^{t_0}  \beta(s,t_0, x_0, \hat{y}_s) \nu(s,t_2) ds. 
\end{equation}
}}
\begin{example}\label{example_ca_future}
Let us to set 
\begin{align*}
g(T) &= \exp(-kT) \\
h(t) &= \sigma
\end{align*}
\end{example}
With this parametrization, the Cheyette model is equiavalent to Hull-White model. It is easy to show that under this parametrization, the convexity adjustment (\ref{ca_approximation_futures}) is
\begin{align*}
CA(t,t_0,t_1) & \approx \frac{P_{E}(0,t_1)}{\delta_{t_1,t_2} P_{E}(0,t_2)} \left(\exp(-k(t_1 - t_0)) - \exp(-k(t_1 - t_0)) \right) \sigma^{2} \exp(-k t_0)  \int_{0}^{t_0} \frac{\left(\exp(- k s) - \exp(- k t_2)   \right)}{k} ds.   
\end{align*}
In the next figure, we can check the accuracy of the last formula versus montecarlo. The parameters that we have used are $\sigma=0.05$, $k=0.003$ and flat curve with level $r=0.01$.

@@

\subsection{OIS futures}
In this we will derive the covexity adjustment for overnight-indexed swap (OIS) under. We will define for $t_0 < t_1$
\begin{align*}
R(t_0,t_1) &= \frac{\mathbb{E}_t^{Q}\left(\exp\left(\int_{t_0}^{t_1}r_{ois,u} du\right)\right) - 1}{t_1 - t_0} \\
R_{avg}(t_0,t_1) &= \frac{\int_{t_0}^{t_1}r_{ois,u} du}{t_1 - t_0}.
\end{align*}
We must observe, that both $R(\cdot,t_0,t_1)$ and $R_{avg}(\cdot,t_0,t_1)$  are not predictible and they are only observable in $t_1$. However, since
$R(\cdot,t_0,t_1)$ and $R_{avg}(\cdot,t_0,t_1)$ are flows that will be payed in $t_1$, we can considerer the expected value under the measure $\mathbb{Q}^{t_1}$ that will observable during whole period $[t_0, t_1]$. Let us define the next $\mathbb{Q}^{t_1}$ martingales  
\begin{align*} 
\bar{R}(t,t_0,t_1) &= \mathbb{E}^{\mathbb{Q}^{t_1}}\left( R(t_0,t_1)  \right) \\
\bar{R}_{avg}(t,t_0,t_1) &= \mathbb{E}^{\mathbb{Q}^{t_1}}\left( R_{avg}(t_0,t_1)  \right).
\end{align*}
We must do several obeservations. The first observation is that
\begin{align*}
F(t,t_0,t_1)=\mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R(t_0,t_1) \right) &= \frac{P_{ois}(t,t_0)- P_{ois}(t,t_1)}{t_1 - t_0}, \quad t \in [0,t_0] \\
F(t,t_0,t_1)=\mathbb{E}_{t}^{\mathbb{Q}}\left(\exp\left(-\int_{t}^{t_0} r_{ois,u} du \right) R(t_0,t_1) \right) &= \frac{1}{t_1 - t_0} \left(\frac{\left( \exp\left(\int_{t}^{t_1}r_{ois,u} du\right)\right)}{P_{ois}(t_0,t)}-1\right), \quad  t \in  [t_0, t_1].
\end{align*}
The can define the convexity adjustment for $ R(t_0,t_1)$ the next way
\begin{equation}\label{R_ois_ca}
CA_{ois}(t,t_0,t_1) = F(t,t_0,t_1) - \bar{R}(t,t_0,t_1) 
\end{equation}
The second observation, is that under the dyam we must do is that
\begin{equation*}
\mathbb{E}_{t}^{\mathbb{Q}}\left(R_{avg}(t_0,t_1) \right) = \mathbb{E}_t^{\mathbb{Q}}\left( \frac{\log\left(1+(t_1 - t_0)R(t_0,t_1) \right)}{t_1 - t_0} \right).  
\end{equation*}
Then, we have two convexity effects, one of them is the convexity of $\log(1+x)$ and the other comes from $R(t_0,t_1)$. Therefore, we need only to compute convexity adjusment for $R(t_0,t_1)$ and apply 
\begin{equation}\label{log_approximation}
\frac{\log(1 + (t_1 - t_0)(x +\epsilon))}{(t_1 - t_0)} \approx \log(1 + (t_1 - t_0)x) + \frac{1}{1 + (t_1 - t_0)x}   \epsilon - \frac{1}{(1 + (t_1 - t_0)x)^{2}} \epsilon^{2}
\end{equation}
to get the convexity adjustment for $R_{avg}(t_0,t_1)$. Now, if we use (\ref{clark-okone}) and (\ref{malliavin_derive_x}, we have the next representation for $R(t_0,t_1)$ 
\begin{align*}
R(t_0,t_1) &=   \mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1) \right) + \int_{0}^{t_1}\mathbb{E}^{\mathbb{Q}}_s\left(D_s R(t_0,t_1) \right) dW^{\mathbb{Q}}_s \\
&= \mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1) \right) + \int_{0}^{t_1}\mathbb{E}^{\mathbb{Q}}_s\left(\int_{\max(s,t_0)}^{t1} D_s x_u du \right) dW^{\mathbb{Q}}_s \\
&= \mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1) \right) + \int_{0}^{t_1}\int_{\max(s,t_0)}^{t1} \mathbb{E}^{\mathbb{Q}}_s\left(\beta(s,u, x_s, y_s)M(s,u)  \right) du  dW^{\mathbb{Q}}_s\\
&\approx \mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1) \right) + \int_{0}^{t_1}\int_{\max(s,t_0)}^{t1} \mathbb{E}^{\mathbb{Q}}_s\left(\beta(s,u, x_s, y_s)M(s,u)  \right) du  dW^{\mathbb{Q}}_s
\end{align*}
Then, if we use Girsanov's theorem we have that
\begin{align*}
F(0,t_0,t_1) &\approx \mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1) \right) - \int_{0}^{t_1}\int_{\max(s,t_0)}^{t1} \mathbb{E}^{\mathbb{Q}}_s\left(\beta(s,u, x_s, y_s)M(s,u)  \right) \nu(s,t_1)  du ds \\
&\approx \mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1) \right) - \int_{0}^{t_1}\int_{\max(s,t_0)}^{t1}  \beta(s,u, x_0, \hat{y}_s) \nu(s,t_1)  du ds  
\end{align*}
And therefore
\begin{equation}\label{convexity_ois_future}
 \mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1) \right) \approx F(0,t_0,t_1) + \int_{0}^{t_1}\int_{\max(s,t_0)}^{t1}  \beta(s,u, x_0, \hat{y}_s) \nu(s,t_1)  du ds.
\end{equation}
Now, if we use (\ref{log_approximation}) with 
\begin{align*}
x &= \mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1) \right)\\
\epsilon &= \int_{0}^{t_1}\int_{\max(s,t_0)}^{t1}  \beta(s,u, x_0, \hat{y}_s) \nu(s,t_1)  du  dW^{\mathbb{Q}}_s
\end{align*} 
we can calculate the adjusment for $R_{avg}(t_0,t1)$
\begin{align*}
\mathbb{E}^{\mathbb{Q}}\left(R_{avg}(t_0,t_1) \right) &\approx \frac{\log\left(1+(t_1 - t_0)\mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1)\right)\right)}{t_1 - t_0} \\
&-\frac{1}{2(1 + (t_1 - t_0)\mathbb{E}^{\mathbb{Q}}\left(R(t_0,t_1)\right))^{2}} \int_{0}^{t_1}\left(\int_{\max(s,t_0)}^{t1}  \beta(s,u, x_0, \hat{y}_s)  du\right)^{2} ds
\end{align*}

\begin{example}
The same way that in the example (\ref{example_ca_future}), we will suppose a Hull-White model with constan mean reversion $k$ and constant volatility $\sigma$. Then
\end{example}
